\documentclass[10pt, oneside]{article} 
\usepackage{amsmath, amsthm, amssymb, calrsfs, wasysym, verbatim, bbm, color, graphics, geometry}


\geometry{tmargin=.75in, bmargin=.75in, lmargin=.75in, rmargin = .75in}  

\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\Cdot}{\boldsymbol{\cdot}}

\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
\newtheorem{conv}{Convention}
\newtheorem{rem}{Remark}
\newtheorem{lem}{Lemma}
\newtheorem{cor}{Corollary}


\title{Mujtaba Siddiqui: Peer evaluation}
\author{[Chirayu Salgarkar]}
\date{Fall 2024}

\begin{document}

\maketitle
\tableofcontents

\vspace{.25in}

\section{CAR T-Cell therapy}
Delivering DNA into cells allows for advanced treatment in disease, e.g. Car-T therapy. They are very expensive. THe hardest part is the transvection process. Technical expertice is hard.


I really like the speed of this presentation. Presenter is clearly a great communicator. 
Image of carbon nanotube arrays allows for tranfsection, which creates vertically aligned carbon tubes to a 13 milimeter AAO wafer. This is key to getting 

About $84\%$ effective. How do you build these devices?

\subsection{Manufacturing}
\begin{itemize}
    \item template AAO
    \item CVD (chemical vapor deposition)
    \item Oxygen plasma etching using ion mill 
    \item RIE or ICBRIE
\end{itemize}

How do you improve the yield?
\subsubsection{CVD}
\begin{itemize}
    \item CVD - carbon deposition using ethylene, high temperatures break down the gas, and carbon sublimes into the deposit.
    \item Variability of this can be done through carbon depsition boat information. Edge cases exist in tube thickness, pore diameter, tip height. There is a corellation between tube thickness and tip height - that is, thicker tubes mean you have taller tip heights.
\end{itemize}
\subsubsection{Reactive Ion Etching}
Don't really understand this ngl

\subsubsection{Intra-device variability}
Chaotic graphs, not a single patterns, either across carrier wafer or across the entire device itself. 

Ultimately, built a processing mechanism to improve yield of these carbon nanotubes, tube thickness improved, thin in general. 







\end{document}
